---
title: "FML Assignment 03"
author: "Sai Supriya"
date: '2022-10-15'
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## \-\--

#Importing the Dataset

```{r}
library(readr)
UniversalBank_1_ <- read_csv("Downloads/UniversalBank (1).csv")
str(UniversalBank_1_)


```

#calling Libraries

```{r}
library(caret)

library(class)

library(ISLR)

```

#Converting Personal Loan Variable

```{r}
UniversalBank_1_$`Personal Loan`=as.factor(UniversalBank_1_$`Personal Loan`)
summary(UniversalBank_1_)
```

#Converting Online Variable

```{r}
UniversalBank_1_$Online = as.factor(UniversalBank_1_$Online) 

summary(UniversalBank_1_$Online)
```

#Converting Creditcard Variable

```{r}
UniversalBank_1_$CreditCard = as.factor(UniversalBank_1_$CreditCard) 
summary(UniversalBank_1_)

UniversalBank_1_$Online<-as.factor(UniversalBank_1_$Online)
is.factor(UniversalBank_1_$Online)
UniversalBank_1_$CreditCard<-as.factor(UniversalBank_1_$CreditCard)
is.factor(UniversalBank_1_$CreditCard)
str(UniversalBank_1_)


```

#Task_A

#Data Partition

```{r}
set.seed(64064)
library(caret)
Train_Index = createDataPartition(UniversalBank_1_$`Personal Loan`,p=0.60, list = FALSE) # 60% reserved for Train
Train.df=UniversalBank_1_[Train_Index,]
Validation.df=UniversalBank_1_[-Train_Index,]

```
##Data Normalization

```{r}
train.norm.df<-Train_Index_1
validation.norm.df<-validation
test.norm.df<-UniversalBank_1_
maindata.norm.df<- UniversalBank_1_

```

```{r}

mytable<- xtabs(~CreditCard+Online+`Personal Loan`, data = Train.df)


ftable(mytable)

```

#Task_B:Consider the task of classifying a customer who owns a bank credit card and is actively using 
online banking services. Looking at the pivot table, what is the probability that this customer 
will accept the loan offer? [This is the probability of loan acceptance (Loan = 1) conditional on 
having a bank credit card (CC = 1) and being an active user of online banking services (Online 
= 1)]. 

```{r}
Probability = 59/(479+59)

Probability

```

#Task_C:

#Create two separate pivot tables for the training data. One will have Loan (rows) as a 
function of Online (columns) and the other will have Loan (rows) as a function of CC. 
```{r}
table(CreditCard=Train.df$CreditCard, `Personal Loan`=Train.df$`Personal Loan`)



```

#pivot table with Personal loan as row and Online as column using training data.

```{r}
table(Online=Train.df$Online, `Personal Loan`=Train.df$`Personal Loan`)


```

#pivot table for Personal loan

```{r}
table(`Personal Loan`=Train.df$`Personal Loan`, CreditCard=Train.df$CreditCard)
```

#Task_D:

#i.P(CC = 1 \| Loan = 1)(the proportion of credit card holders among the loan acceptors)

```{r}
Probability_1 = 93/(195+93)

Probability_1

```

#ii.P(Online = 1 \| Loan = 1)

```{r}
Probability_2 = 179/(109+179)

Probability_2

```

#iii.P(Loan = 1) (the proportion of loan acceptors)

```{r}
Probability_3 = 288/(2712+288)

Probability_3

```

#iv.P(CC = 1 \| Loan = 0)

```{r}
Probability_4 = 788/(1924+788)

Probability_4

```

#v.P(Online = 1 \| Loan = 0)

```{r}
Probability_5 = 1631/(1631+1081)

Probability_5
```

#vi.P(Loan = 0)

```{r}
Probability_6 = 2712/(2712+288)

Probability_6
```

#Task_E:Use the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC 
= 1, Online = 1). 

#P(Loan = 1 \| CC = 1, Online = 1).

```{r}
naive_Bayes_probability <- (Probability_1*Probability_2*Probability_3) / 

                          ((Probability_1*Probability_2*Probability_3) + 

                             (Probability_4*Probability_5*Probability_6))


naive_Bayes_probability
```

#Task_F:Compare this value with the one obtained from the pivot table in (B).

#Which is a more accurate estimate?

```{r}
#0.1087106 in task-E is very similar to the  0.1096654 in task-B.

#The difference between the exact and naive bayes methods is that

#the exact approach requires the same independent variable classifications to predict, 

#whereas the naive bayes method does not.



```

#Task_G:

\# P(Loan = 1 \| CC = 1, Online = 1)

#Run naive Bayes on the data. Examine the model output on training data, and find the entry

#that corresponds to P(Loan = 1 \| CC = 1, Online = 1). Compare this to the number you
#obtained in (E).

```{r}
library(e1071)

```

```{r}

library(naivebayes)
library(mlbench)
nb.model<- naiveBayes(`Personal Loan`~Online+CreditCard, data= Train.df)

To_Predict=data.frame(Online= '1', CreditCard= '1')

predict(nb.model,To_Predict,type='raw')


```

#The task-G value of 0.1087106 and the task-E value of 0.1087106 are identical.

#As a result, the naive bayes produces the same results as the prior approaches.
